---

GenomicRanges_Rle_Note
---

```{r dependencies, warning=FALSE, message=FALSE}
library(GenomicRanges)
```

### Overview

`Rle` (run length encoding) is a class great for representation genome-wide sequence coverage.

### Coverage

In high-throughput sequencing, coverage is the number of reads overlapping each base. 

This is a fundamental quantity for many high-throughout sequencing analyses. For variant calling (DNA sequencing) it tells you how much power (information) you have to call a variant at a given location.  For ChIP sequencing it is the primary signal; areas with high coverage are thought to be enriched for a given protein.

### Rle

An `Rle` (run-length-encoded) vector is a specific representation of a vector.  The `Biocpkg("IRanges")` package implements support for this class.

```{r RleEx1}
rl <- Rle(c(1,1,1,1,2,2,3,3,2,2))
rl

# accessor functions `runLength()` and `runValue()`
runLength(rl)
runValue(rl)
as.numeric(rl)
```

This is a very efficient representation if

- the vector is very long
- there are a lot of consecutive elements with the same value

This is especially useful for genomic data which is either piecewise constant, or where most of the genome is not covered (eg. RNA sequencing in mammals).

You can do arithmetic with them, transform them etc. using standard R functions like `+` and `log2`.

There are also `RleList` which is a list of `Rle`s.  This class is used to represent a genome wide coverage track where each element of the list is a different chromosome.

### Useful functions for Rle

A standard usecase is that you have a number of regions (say `IRanges`) and you want to do something to your `Rle` over each of these regions.  Enter `aggregate()`.

```{r aggregate}
ir <- IRanges(start = c(2,6), width = 2)
ir
aggregate(rl, ir, FUN = mean)

```

It is also possible to covert an `IRanges` to a `Rle` by the `coverage()` function.  This counts, for each integer, how many ranges overlap the integer.
```{r coverage}
ir <- IRanges(start = 1:10, width = 3)
ir
rl <- coverage(ir)
rl
```

You can select high coverage regions by the `slice()` function:

```{r slice}
slice(rl, 1)
slice(rl, 2)
```

This outputs a `Views` object.

### Views and Rles

We learned about `Views` on genomes.  `Views` can also be instantiated on `Rle`s.

```{r views}
vi <- Views(rl, start = c(3,7), width = 3)
vi
``` 

with `Views` you can now (again) apply functions:
```{r viewsMean}
mean(vi)
```
This is very similar to using `aggregate()` described above.

### RleList

An `RleList` is simply a list of `Rle`.  It is similar to a `GRangesList` in concept.

### Rles and GRanges

`Rle`'s can also be constructed from `GRanges`.

This often involves `RleList` where each element of the list is a chromosome.  Surprisingly, we do not yet have an `RleList` type structure which also contains information about say the length of the different chromosomes.


```{r GRanges}
gr <- GRanges(seqnames = "chr1", ranges = IRanges(start = 1:10, width = 3))
gr
rl <- coverage(gr)
rl
```

Using `Views` on such an object exposes some missing functionality

```{r GRangesViews, error=TRUE}
grView <- GRanges("chr1", ranges = IRanges(start = 2, end = 7))
grView
vi <- Views(rl, grView)
vi
```

We get an error, mentioning some object called a `RangesList`.

```{r eval = FALSE}
vi <- Views(rl, as(grView, "RangesList"))
vi
vi[[1]]
```

## Biology Usecase

Suppose we want to compute the average coverage of bases belonging to (known) exons.

Input objects are   
**reads**: a `GRanges`.  
**exons**: a `GRanges`.

pseudocode:
```{r usecase, eval=FALSE}
bases <- reduce(exons)
nBases <- sum(width(bases))
nCoverage <- sum(Views(coverage(reads), bases))
nCoverage / nBases
``` 

(watch out for strand)
