---
Di Zhen

GenomicRanges_GRanges_Note
---

```{r dependencies, warning=FALSE, message=FALSE}
library(GenomicRanges)
```

## GRanges

`GRanges` are like `IRanges` with strand and chromosome.  Strand can be `+`, `-` and `*`. The value `*` indicates 'unknown strand' or 'unstranded'. 
```{r GRanges}
# create a GRange
gr <- GRanges(seqnames = "chr1", strand = c("+", "-", "+"),
              ranges = IRanges(start = c(1,3,5), width = 3))
gr              
``` 

Natural accessor functions: `strand()`, `seqnames()`, `ranges()`, `start()`, `end()`, `width()`.

Operations relative to the direction of transcription (`upstream()`, `downstream()`):

```{r flank}
flank(gr, 2, start = FALSE) # strand = "-": upstream; stand = "+": downstream
``` 


## GRanges, seqinfo

`GRanges` operate within a universe of sequences (chromosomes/contigs) and their lengths.

This is described through `seqinfo`:

```{r seqinfo}
seqinfo(gr)
seqlengths(gr) <- c("chr1" = 10)
seqinfo(gr)
seqlevels(gr)
seqlengths(gr)
``` 

Especially the length of the chromosomes are used by some functions.  For example `gaps()` return the stretches of the genome not covered by the `GRanges`.

```{r gaps}
gr
gaps(gr)
```

In this example, we know that the last gap stops at 10, because that is the length of the chromosome. Note how a range on the `*` strand appears in the result.

Let us expand the `GRanges` with another chromosome

```{r gr2}
seqlevels(gr) <- c("chr1", "chr2")
seqnames(gr) <- c("chr1", "chr2", "chr1")
gr
```

When you `sort()` a `GRanges`, the sorting order of the chromosomes is determined by their order in `seqlevel`. 

```{r sort}
sort(gr)
seqlevels(gr) <- c("chr2", "chr1")
sort(gr)
```

associate a genome with a `GRanges`. 
```{r genome}
genome(gr) <- "hg19"
gr
```

It allows R to throw an error when you compare two `GRanges` from different genomes, like

```{r gr-error, error=TRUE}
gr2 <- gr
genome(gr2) <- "hg18"
findOverlaps(gr, gr2)
```
